### 2022-10-03 Milestone2.Rmd
Deleting excel files not pertinent

Milestone 2

```{r, echo=FALSE}
library(readxl)
library(tidyr)
library(lubridate)

```

```{r}
MPX_data <- read.csv("~/251/files/euro_mpx_cases.csv")
```

•	Description of dataset
o	What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
o	How does the dataset relate to the group problem statement and question?

/newpage

•	Import statement
o	NOTE: Please use datasets available in the PHW251 Project Data github repo Links to an external site. (this is important to make sure everyone is using the same datasets)
o	Use appropriate import function and package based on the type of file
o	Utilize function arguments to control relevant components (i.e. change column types, column names, missing values, etc.)
o	Document the import process

/newpage

•	Identify data types for 5+ data elements/columns/variables
o	Identify 5+ data elements required for your specified scenario. If <5 elements are required to complete the analysis, please choose additional variables of interest in the data set to explore in this milestone.
o	Utilize functions or resources in RStudio to determine the types of each data element (i.e. character, numeric, factor)
o	Identify the desired type/format for each variable—will you need to convert any columns to numeric or another type?

/newpage

•	Provide a basic description of the 5+ data elements
o	Numeric: mean, median, range
o	Character:  unique values/categories
o	Or any other descriptives that will be useful to the analysis

### 2022-10-03 Milestone2.Rmd
Worked on Milestone2 together

@@
-Milestone 2
-
+---
+title: "Milestone 2"
+names: "Dawson Coblin & Jose Lazo"
+date: "October 3, 2022"
+output: 
+---
 ```{r, echo=FALSE}
@@
 
-•	Description of dataset
+Description of dataset
+
 o	What is the data source? (1-2 sentences on where the data is coming from, dates included, etc.)
+The monkeypox data file contains information on the number of monkeypox cases reported by EU/EEA countries or collected throughout epidemiologic intelligence at ECDC. 
+Each row contains the corresponding data for a country, day of reporting, number of cases and source of information (data are in long format).
+The file is updated twice a week.
+The date range for this data set is from 05/09/22 - 08/23/22
+
 o	How does the dataset relate to the group problem statement and question?
+For our project we are interested in understanding how case rates may differ by region and various demographic factors. The MPX data has information on confirmed cases by countries.
 
@@
 
+Import process
+1) Download zip file from phw251_projectdata
+2) Set up our github repository on our R datahub
+3) One team member uploaded pertinent excel files to their Rproject 251
+4) Commit and push to have excel files on our team's github
+
 /newpage
@@
 
+```{r}
+str(MPX_data)
+mean(MPX_data$ConfCases)
+```
 /newpage
@@
 o	Or any other descriptives that will be useful to the analysis
+
+Confcases is a numeric data element
+```{r}
+round(mean(MPX_data$ConfCases),2)
+median(MPX_data$ConfCases)
+range(MPX_data$ConfCases)
+```
+Mean = 5.72
+Median = 0
+Range = 0,655
+
+DateRep 
+Is a character variable which contains dates 05/09/22 through 08/23/22. We will convert this data into numeric form.
+
+```{r}
+unique(MPX_data$CountryExp)
+```
+
+CountryExp
+Is a character variable which contains 29 countries listed above.
+
+```{r}
+unique(MPX_data$CountryCode)
+```
+
+CountryCode
+Is a character with the 2 letter abbreviation of countries with 29 values which matches with CountryExp
+
+```{r}
+unique(MPX_data$Source)
+```
+
+Source is a character variable with 2 options, TESSy and EI

### 2022-11-05 MS3.Rmd
Add files via upload

---
title: "Scenario 2 - MPX in the EU"
author: "Dawson Coblin and Jose Lazo"
date: '2022-11-07'
output: pdf_document
---

# Monkeypox (MPX) Infection Rates in the European Union

This report aims to prepare data to analyze and assess the impact of MPV within  
each country of the EU, and understand how case rates may differ by region and 
other various demographic factors. These findings will aid the implementation
and appropriate response to the epidemic to local health departments.

### Legend of Tasks and Milestones

Milestone  | Description of Section
---------- | ----------------------
Milestone 1|Group Agreement
Milestone 2|Data-set Imported into R; Identify Key Data Elements and Types
Milestone 3|Clean Data-set Create Descriptive Statistics and Data Dictionary

Milestone 3 Sub-sections:  
1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames  
2.Summarizing Data and Creating New Vectors, Organize Vectors for Use  
3.Clean Variables and Combined Data-sets  
4.Descriptive Stats of EU Data and Data Dictionary

\newpage

### 1) Milestone 1: Group Agreement

\newpage
### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types

```{r packages}
library(ggplot2)
library(tidyverse)
library(readr)
library(lubridate)
library(stringr)
library(readxl)
```
```{r}
euro_mpx_cases <- read_csv("files/euro_mpx_cases.csv")
euro_census_stats <- read_csv("files/euro_census_stats.csv")
euro_pop_denominators <- read_csv("files/euro_pop_denominators.csv")
world_country_regions <- read_csv("files/phw251_projectdata-d6f7578ccda8b47b98f4ab30005bbd76acf496ec/world_country_regions.csv")
```
\newpage

### Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames

Data Frame Cleaning

```{r}
#Cleaning the cases df
euro_mpx_cases <- euro_mpx_cases %>%
  select( DateRep, CountryExp, CountryCode  ,ConfCases)

#Creating a vector to use for country code/country name and regions in EU
euro_country_vector <- unique(euro_mpx_cases$CountryCode)

#Cleaning the denominators df
euro_pop_denominators <- euro_pop_denominators %>% 
  filter( TIME_PERIOD == 2022) %>%
  rename( CountryCode = geo , total_pop_2022 = 'OBS_VALUE') %>%
  select( CountryCode, total_pop_2022)

euro_pop_denominators <- euro_pop_denominators[euro_pop_denominators$CountryCode 
                                               %in% euro_country_vector, ]

#Clean census data to include only country code, res_pop. The date for the census 
#dataframe is 2011, so the date is removed and not relevant.

euro_census_stats <- euro_census_stats %>%
  group_by(COUNTRY_CODE, RES_POP) %>%
  summarise(census_pop = sum(pop)) %>%
  rename(CountryCode = COUNTRY_CODE)

#Widen census info to allow each country row to have pop information.
euro_census_stats <- 
  pivot_wider(euro_census_stats, names_from = RES_POP, values_from = census_pop)

euro_census_stats$`0-1000`[euro_census_stats$`0-1000` == 0] <- NA
euro_census_stats$`1000-9999`[euro_census_stats$`1000-9999` == 0] <- NA
euro_census_stats$`10000-99999`[euro_census_stats$`10000-99999` == 0] <- NA
euro_census_stats$`100000-199999`[euro_census_stats$`100000-199999` == 0] <- NA
euro_census_stats$`200000-499999`[euro_census_stats$`200000-499999` == 0] <- NA
euro_census_stats$`500000-999999`[euro_census_stats$`500000-999999` == 0] <- NA
euro_census_stats$GE1000000[euro_census_stats$GE1000000 == 0] <- NA
```

\newpage

### Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use

```{r}
#We then assign counties with a region code in a new column a list "EU_region,"
#and include the country name and code
sr_northern_e<- c(
  "Denmark" , "Estonia", "Finland", "Iceland", "Ireland", "Latvia", "Lithuania",
  "Norway", "Sweden" , "DK" , "EE", "FI", "IE", "IS", "LT", "LU", "NO", "SE"
)
sr_western_e<- c(
  "Austria", "Belgium", "France", "Germany", "Luxembourg", "Netherlands" , 
  "DE", "AT" , "BE" , "FR", "LV", "NL"
)
sr_eastern_e<- c(
  "Bulgaria", "Czechia", "Hungary", "Poland", "Romania", "Slovakia", "BG", "CZ",
  "HU", "PL" , "RO", "SK"
)
sr_southern_e<- c(
  "Croatia", "Greece", "Italy", "Malta", "Portugal", "Slovenia", "Spain", "EL", 
  "ES", "HR", "IT", "MT", "PT", "SI"
)
sr_western_a<- c( 
  "Cyprus" , "CY"
)

#Create a floor month for the date range to aggregate by month
euro_mpx_cases <- euro_mpx_cases %>%
  mutate(floor_month = floor_date( DateRep , "month"))

#Review the aggregate totals as a value of confirmed cases in Europe and list as
#"cases_grouped_monthly"

euro_mpx_cases <- euro_mpx_cases %>% 
  group_by(floor_month, CountryCode) %>% 
  summarize(cases_grouped_monthly = sum(ConfCases))

#We now add the EU_Region vector based on the country name.
euro_mpx_cases <- euro_mpx_cases %>% 
  mutate(EU_region = case_when(
    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
    CountryCode %in% sr_northern_e ~ "Northern Europe",
    CountryCode %in% sr_southern_e ~ "Southern Europe",
    CountryCode %in% sr_western_e ~ "Western Europe",
    CountryCode %in% sr_western_a ~ "Western Asia"
    ))
```

\newpage

### Milestone 3: 3. Clean Variables and Combined Data-sets

```{r}

#Combined the denominations data for 2022 pop with euro_mpx_cases dataset

euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
                                by= "CountryCode")
#Join census data
euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
                              by= "CountryCode")
#Rearrange column order
euro_mpx_cases <- relocate(euro_mpx_cases, floor_month, EU_region, 
                          total_pop_2022 , cases_grouped_monthly, 
                          .before = CountryCode)

euro_mpx_cases<- relocate(euro_mpx_cases, CountryCode, .before = total_pop_2022)

```

\newpage

### Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary

Descriptive stats:
We find that the census information for 7 of the 29 countries in the EU are not
available. These "0" values were replaced with "NA" to prevent graphing issues.

Region Data:
We see that the regions in EU include various numbers of countries, thus having unequal
populations.

"Eastern Europe" : 6 countries, total region population = 89171711   
"Northern Europe": 9 countries, total region population = 37518701   
"Southern Europe": 7 countries, total region population = 133879004   
"Western Asia"   : 1 countries, total region population = 904705   
"Western Europe" : 6 countries, total region population = 190251495

```{r}

visual <- ggplot(data = euro_mpx_cases, 
                 aes(x = EU_region , y = cases_grouped_monthly))+ 
  theme_minimal(base_size = 15)

visual

```

```{r}
#Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
Data_Dictionary

```

### 2022-11-06 MS3.Rmd
working on data dictionary

@@
 euro_pop_denominators <- read_csv("files/euro_pop_denominators.csv")
-world_country_regions <- read_csv("files/phw251_projectdata-d6f7578ccda8b47b98f4ab30005bbd76acf496ec/world_country_regions.csv")
+world_country_regions <- read_csv("files/world_country_regions.csv")
 ```
@@
 #Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
-Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
-Data_Dictionary
+#Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
+#Data_Dictionary
+
+Floor month
+year-MM-DD
+
+EU region (European Region)
+Western Europe
+Eastern Europe
+Western Europe
+Northern Europe
+Souther Europe
+
+Country Code (2 Letter Alpha Country Code)
+AT Austria
+BE Belgium
+BG Bulgaria
+CY Cyprus
+CZ Czechia
+DE Germany
+DK Denmark
+EE Estonia
+EL Greece
+ES Spain
+FI Finland
+FR France
+HR Croatia
+HU Hungary
+IE Ireland
+IS Iceland
+IT Italy
+LT Lithuania
+LU Luxembourg
+LV Latvia
+MT Malta
+NL Netherlands
+NO Norway
+PL Poland
+PT Portugal
+RO Romania
+SE Sweden
+SI Slovenia
+SK Slovakia
+
+total_pop_2022
+
+cases_grouped monthly (Monkeypox cases per month in each country)
+
+0-1000
+
+1000-9999
+
+10000-99999
+
+100000-199999
+
+200000-499999
+
+500000-999999
+
+GE1000000
 

### 2022-11-07 MS3.Rmd
Updates

@@
 
-Floor month
-year-MM-DD
+```
 
-EU region (European Region)
-Western Europe
-Eastern Europe
-Western Europe
-Northern Europe
-Souther Europe
+|Floor month - Date|
+|:----------------:|
+|year-MM-DD|
+|:----------------:|
+|EU region (European Region)|
+|:----------------:|
+|Western Europe|
+|Eastern Europe|
+|Western Europe|
+|Northern Europe|
+|Souther Europe|
 
-Country Code (2 Letter Alpha Country Code)
-AT Austria
-BE Belgium
+##Country Code (2 Letter Alpha Country Code)
+###AT Austria
+###BE Belgium
 BG Bulgaria
@@
 GE1000000
-
-```

### 2022-11-07 MS3.Rmd
11/6/22 Upload

--- 2022-11-05 MS3.Rmd
@@
 
+Group R: Team Logistics Agreement
+
+9/19/22
+
+1.	Project option selection
+
+Scenario 2: Monkeypox (MPX) infection rates in the European Union
+
+2.	Create git repository
+
+https://github.com/dcoblin/251.git
+
+3.	Roles and responsibilities
+
+a.	Team's preferred communication method
+
+WhatsApp
+
+b.	Team's preferred meeting times and frequency
+
+Weekends/Post 6 PM on weekdays
+
+c.	Team's preferred method for tracking progress
+
+Google docs
+
+d.	Point person for contacting course facilitators with questions
+
+Jose/Dawson
+
+e.	Point person for submitting all milestones
+
+Dawson/Jose
+
 \newpage
@@
 
-### Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames
+## Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames
 
-Data Frame Cleaning
+### Data Frame Cleaning
 
@@
   filter( TIME_PERIOD == 2022) %>%
-  rename( CountryCode = geo , total_pop_2022 = 'OBS_VALUE') %>%
-  select( CountryCode, total_pop_2022)
+  rename( CountryCode = geo , country_pop_2022 = 'OBS_VALUE') %>%
+  select( CountryCode, country_pop_2022)
 
@@
 
-### Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use
+## Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use
 
@@
     ))
-```
-
-\newpage
-
-### Milestone 3: 3. Clean Variables and Combined Data-sets
-
-```{r}
-
-#Combined the denominations data for 2022 pop with euro_mpx_cases dataset
-
-euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
-                                by= "CountryCode")
-#Join census data
-euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
-                              by= "CountryCode")
-#Rearrange column order
-euro_mpx_cases <- relocate(euro_mpx_cases, floor_month, EU_region, 
-                          total_pop_2022 , cases_grouped_monthly, 
-                          .before = CountryCode)
-
-euro_mpx_cases<- relocate(euro_mpx_cases, CountryCode, .before = total_pop_2022)
 
@@
 
-### Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary
+## Milestone 3: 3. Clean Variables and Combined Data-sets
 
-Descriptive stats:
+```{r}
+
+euro_mpx_cases <-  inner_join( euro_pop_denominators, euro_mpx_cases, 
+                                by= "CountryCode")
+#Join census data.
+euro_mpx_cases <- inner_join(euro_census_stats, euro_mpx_cases,
+                              by= "CountryCode")
+#Rearrange column order.
+euro_mpx_cases <- relocate(euro_mpx_cases, EU_region, floor_month, 
+                          country_pop_2022 , cases_grouped_monthly, 
+                          .before = CountryCode)
+
+#Combined the denominations data for 2022 pop with euro_mpx_cases dataset.
+sum_region_cases <- euro_mpx_cases %>% group_by(EU_region) %>%
+  summarise(sum(cases_grouped_monthly))
+
+#Join the numeric sums into the main dataframe.
+euro_mpx_cases <- inner_join(sum_region_cases, euro_mpx_cases,
+                              by= "EU_region")
+
+#Relocate the columns to a adjust fluidity of the data.
+euro_mpx_cases<- relocate(euro_mpx_cases, floor_month, CountryCode, 
+                          country_pop_2022, cases_grouped_monthly, 
+                          .before = EU_region)
+#Rename the sum of region cases that are confirmed.
+euro_mpx_cases<- euro_mpx_cases %>%
+   rename( sum_region_cases = `sum(cases_grouped_monthly)`)
+```
+
+\newpage
+
+## Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary
+
+### Descriptive stats:
 We find that the census information for 7 of the 29 countries in the EU are not
@@
 
-Region Data:
-We see that the regions in EU include various numbers of countries, thus having unequal
-populations.
+### Region Data:
+We see that the regions in EU include various numbers of countries, thus having 
+unequal populations. A useful metric to create a graph will be the calculation 
+of cases per 100,000. This will allow us to use 1 categorical variable (EU Region)
+to a quantitative variable (number of confirmed cases)
 
-"Eastern Europe" : 6 countries, total region population = 89171711   
-"Northern Europe": 9 countries, total region population = 37518701   
-"Southern Europe": 7 countries, total region population = 133879004   
-"Western Asia"   : 1 countries, total region population = 904705   
-"Western Europe" : 6 countries, total region population = 190251495
+We compiled the relevant confirmed case information per region in the EU.
 
-```{r}
+### The pertinent information per EU region is listed below:
+```{r echo=FALSE}
+Table_1 <- tibble(
+  EU_Region = c("Eastern Europe", "Northern Europe", "Southern Europe",
+               "Western Asia", "Western Europe" ),  
+  Countries_per_Region = c(6, 9, 7, 1, 6),
+  Total_Region_Pop = c(89171711, 37518701,133879004, 904705, 190251495),
+  Total_cases_per_region = c( 217, 619, 7958, 4, 8214)
+  )
 
-visual <- ggplot(data = euro_mpx_cases, 
-                 aes(x = EU_region , y = cases_grouped_monthly))+ 
-  theme_minimal(base_size = 15)
-
-visual
-
+Table_1
 ```
+## Data Dictionary
 
@@
 Data_Dictionary
-
 ```

### 2022-11-07 MS3.Rmd
Created data dictionary

@@
 library(readxl)
+library(kableExtra)
 ```
@@
 ```{r}
-#Upload the Data Dictionary for the central data-frame: "euro_mpx_cases"
-Data_Dictionary <- read_excel("~/PHW251_Fall2022/251/MS3_data_dictionary.xlsx")
-Data_Dictionary
+#Create the Data Dictionary
+
+names <- names(euro_mpx_cases)
+data_type <- sapply(euro_mpx_cases,class)
+description <- c("Reporting time period with the floor month", #floor_month
+                 "Abbreviation of Country", # CountryCode
+                 "The current population of the country for 2022", # country_pop_2022
+                 "Total confirmed MPX cases grouped by month", # cases_grouped_monthly
+                 "The designated region the country is located", # EU_region
+                 "Sum of the four months per country code", # sum_region_cases
+                 "The Sum of population 0-1000", # 0-1000
+                 "The Sum of population 1000-9999", # 1000-9999
+                 "The Sum of population 10000-99999", # 10000-99999
+                 "The Sum of population 100000-199999", # 100000-199999
+                 "The Sum of population 200000-499999", # 200000-499999
+                 "The Sum of population 500000-999999", # 500000-999999
+                 "The Sum of population GE1000000" # GE1000000
+)
+
+data_dic <- as.data.frame(cbind(names, data_type, description)) %>% 
+  select(-names)
+
+kable(data_dic,
+      col.names = c("Data Type", 
+                    "Description"),
+      caption = "Data Dictionary"
+      ) 
 ```

### 2022-11-07 MS3.Rmd
Included a visualization however have hidden it as not needed for MS 3

@@
 ```
+```{r, include=FALSE}
+euro_mpx_cases %>% 
+  ggplot(., aes(x = EU_region, y = country_pop_2022, fill = EU_region)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  theme_minimal() +
+  labs(title = "Total Population 2022 by Region", x = "", y = "Pop") +
+  theme(axis.title = element_text(size = 14),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        legend.title = element_blank(),
+        legend.position = "none")
+```

### 2022-11-07 MS3.Rmd
Test

--- 2022-11-07 MS3.Rmd
@@
 ```
+```{r, include=FALSE}
+euro_mpx_cases %>% 
+  ggplot(., aes(x = EU_region, y = country_pop_2022, fill = EU_region)) +
+  geom_bar(stat = "identity", position = position_dodge()) +
+  theme_minimal() +
+  labs(title = "Total Population 2022 by Region", x = "", y = "Pop") +
+  theme(axis.title = element_text(size = 14),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        legend.title = element_blank(),
+        legend.position = "none")
+```
+TEST

### 2022-11-07 MS3.Rmd
Finalized table and data dictionary and uploading pdf

--- 2022-11-07 MS3.Rmd
@@
 output: pdf_document
+editor_options: 
+  markdown: 
+    wrap: 72
 ---
@@
 
-This report aims to prepare data to analyze and assess the impact of MPV within  
-each country of the EU, and understand how case rates may differ by region and 
-other various demographic factors. These findings will aid the implementation
-and appropriate response to the epidemic to local health departments.
+This report aims to prepare data to analyze and assess the impact of MPV
+within\
+each country of the EU, and understand how case rates may differ by
+region and other various demographic factors. These findings will aid
+the implementation and appropriate response to the epidemic to local
+health departments.
 
@@
 
-Milestone  | Description of Section
----------- | ----------------------
-Milestone 1|Group Agreement
-Milestone 2|Data-set Imported into R; Identify Key Data Elements and Types
-Milestone 3|Clean Data-set Create Descriptive Statistics and Data Dictionary
+| Milestone   | Description of Section                                           |
+|-----------------------|-------------------------------------------------|
+| Milestone 1 | Group Agreement                                                  |
+| Milestone 2 | Data-set Imported into R; Identify Key Data Elements and Types   |
+| Milestone 3 | Clean Data-set Create Descriptive Statistics and Data Dictionary |
 
-Milestone 3 Sub-sections:  
-1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames  
-2.Summarizing Data and Creating New Vectors, Organize Vectors for Use  
-3.Clean Variables and Combined Data-sets  
+Milestone 3 Sub-sections:\
+1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames\
+2.Summarizing Data and Creating New Vectors, Organize Vectors for Use\
+3.Clean Variables and Combined Data-sets\
 4.Descriptive Stats of EU Data and Data Dictionary
@@
 
-https://github.com/dcoblin/251.git
+<https://github.com/dcoblin/251.git>
 
@@
 
+```{=html}
+<!-- -->
+```
 a.  Team's preferred communication method
@@
 \newpage
+
 ### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types
 
-```{r packages}
+```{r packages, include=FALSE}
 library(ggplot2)
@@
 ```
+
 ```{r}
@@
 ```
+
 \newpage
@@
 
+head(euro_pop_denominators,3)
+
 #Clean census data to include only country code, res_pop. The date for the census 
@@
 euro_census_stats$GE1000000[euro_census_stats$GE1000000 == 0] <- NA
+
+head(euro_census_stats, 3)
 ```
@@
 
+head(euro_mpx_cases,3)
 ```
@@
    rename( sum_region_cases = `sum(cases_grouped_monthly)`)
+
+head(euro_mpx_cases,3)
 ```
@@
 ### Descriptive stats:
-We find that the census information for 7 of the 29 countries in the EU are not
-available. These "0" values were replaced with "NA" to prevent graphing issues.
+
+We find that the census information for 7 of the 29 countries in the EU
+are not available. These "0" values were replaced with "NA" to prevent
+graphing issues.
 
 ### Region Data:
-We see that the regions in EU include various numbers of countries, thus having 
-unequal populations. A useful metric to create a graph will be the calculation 
-of cases per 100,000. This will allow us to use 1 categorical variable (EU Region)
-to a quantitative variable (number of confirmed cases)
 
-We compiled the relevant confirmed case information per region in the EU.
+We see that the regions in EU include various numbers of countries, thus
+having unequal populations. A useful metric to create a graph will be
+the calculation of cases per 100,000. This will allow us to use 1
+categorical variable (EU Region) to a quantitative variable (number of
+confirmed cases)
+
+We compiled the relevant confirmed case information per region in the
+EU.
 
 ### The pertinent information per EU region is listed below:
-```{r echo=FALSE}
-Table_1 <- tibble(
-  EU_Region = c("Eastern Europe", "Northern Europe", "Southern Europe",
-               "Western Asia", "Western Europe" ),  
-  Countries_per_Region = c(6, 9, 7, 1, 6),
-  Total_Region_Pop = c(89171711, 37518701,133879004, 904705, 190251495),
-  Total_cases_per_region = c( 217, 619, 7958, 4, 8214)
-  )
 
-Table_1
+```{r, include=FALSE}
+
+Table_1.1 <- euro_mpx_cases %>% group_by(EU_region, country_pop_2022, sum_region_cases) %>%
+  count(CountryCode) %>%
+  select(CountryCode,EU_region,country_pop_2022,sum_region_cases) %>%
+  ungroup() %>%
+  group_by(EU_region, sum_region_cases) %>%
+  summarize(Total_region_pop = sum(country_pop_2022))
+Countries_per_Region = c(6, 9, 7, 1, 6)
+Table_1.1 <- as.data.frame(cbind(Table_1.1, Countries_per_Region))
 ```
+
+```{r, echo=FALSE}
+kable(Table_1.1,
+      col.names = c("EU Region", 
+                    "Total Cases Per Region",
+                    "Total Region Population",
+                    "Countries Per Region"),
+      caption = "Monkeypox Cases Per Region"
+      ) %>%
+  kable_styling(latex_options = "HOLD_position")
+
+```
+
 ## Data Dictionary
 
-```{r}
+```{r, echo=FALSE}
 #Create the Data Dictionary
@@
       caption = "Data Dictionary"
-      ) 
+      ) %>%
+  kable_styling(latex_options = "HOLD_position")
 ```
+
+```{r, include=FALSE}
+#euro_mpx_cases %>% 
+#  ggplot(., aes(x = EU_region, y = country_pop_2022, fill = EU_region)) +
+#  geom_bar(stat = "identity", position = position_dodge()) +
+#  theme_minimal() +
+#  labs(title = "Total Population 2022 by Region", x = "", y = "Pop") +
+#  theme(axis.title = element_text(size = 14),
+#        axis.text = element_text(size = 8),
+#        axis.ticks.x = element_blank(),
+#        legend.title = element_blank(),
+#        legend.position = "none")
+```

### 2022-11-17 MS4.Rmd
Created bar graph and line plot

--- 2022-11-07 MS3.Rmd
@@
 author: "Dawson Coblin and Jose Lazo"
-date: '2022-11-07'
+date: '2022-11-21'
 output: pdf_document
@@
 | Milestone 3 | Clean Data-set Create Descriptive Statistics and Data Dictionary |
+| Milestone 4 | Data Visualization                                               |
 
@@
 
-```{r, include=FALSE}
-#euro_mpx_cases %>% 
-#  ggplot(., aes(x = EU_region, y = country_pop_2022, fill = EU_region)) +
-#  geom_bar(stat = "identity", position = position_dodge()) +
-#  theme_minimal() +
-#  labs(title = "Total Population 2022 by Region", x = "", y = "Pop") +
-#  theme(axis.title = element_text(size = 14),
-#        axis.text = element_text(size = 8),
-#        axis.ticks.x = element_blank(),
-#        legend.title = element_blank(),
-#        legend.position = "none")
+\newpage
+
+### Milestone 4: Visualizations
+
+```{r}
+euro_mpx_cases %>% 
+  ggplot(., aes(x = floor_month, y = cases_grouped_monthly)) +
+  geom_bar(aes(fill = EU_region), stat = "identity") +
+  theme_minimal() +
+  labs(title = "Monthly Cases of Monkeypox in Europe", x = "Month (2022)", 
+       y = "Monkeypox Cases") +
+  theme(axis.title = element_text(size = 10),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        plot.title = element_text(size = 14),
+        legend.title = element_blank()) 
 ```
+
+### Bar Graph
+In the bar graph above we see the breakdown of monthly Monkeypox infections in Europe between May and August of this year.
+The majority of cases occured in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
+Over the course of 4 months, the total number of Monkeypox cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214
+
+\newpage
+
+```{r}
+euro_mpx_cases <- inner_join(euro_mpx_cases, Table_1.1, by= "EU_region") %>%
+  group_by(floor_month) %>%
+  mutate(Total_Euro_cases_monthly = sum(cases_grouped_monthly)) %>%
+  mutate(Total_euro_pop = sum(country_pop_2022))
+
+graph1 <- euro_mpx_cases %>%
+  select(floor_month, cases_grouped_monthly, EU_region, sum_region_cases.x, Total_region_pop,
+         Total_Euro_cases_monthly, Total_euro_pop) %>%
+  group_by(floor_month, EU_region) %>%
+  mutate(region_monthly_cases = sum(cases_grouped_monthly)) %>% ungroup() %>%
+  mutate("monthly_rate_region" = round((region_monthly_cases/Total_region_pop * 100000),2))
+
+ggplot(data=graph1, aes(x=floor_month, y=monthly_rate_region, group= EU_region)) +
+  geom_line(aes(color=EU_region)) +
+  geom_point() +
+  geom_label(label= graph1$monthly_rate_region, nudge_y = 0.1) +
+  theme_minimal() +
+  labs(title = "Rate of Monkeypox Infections", x = "Month (2022)", 
+       y = "Monkeypox Rate (per 100k)") +
+    theme(axis.title = element_text(size = 10),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        plot.title = element_text(size = 14),
+        legend.title = element_blank()) 
+```
+
+### Line Plot
+As a follow up to the bar graph above, we wanted to look at the rate of Monkeypox infections per region.
+In the bar graph above, Western Europe had the greatest number of Monkeypox infections however when we look at the rate of infections, we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.

### 2022-11-21 MS4.Rmd
Adding figure caption to figures

@@
   theme_minimal() +
-  labs(title = "Monthly Cases of Monkeypox in Europe", x = "Month (2022)", 
+  labs(title = "Monthly Cases of Monkeypox in Europe", 
+       tag = "Figure 1",
+       x = "Month (2022)", 
        y = "Monkeypox Cases") +
@@
         plot.title = element_text(size = 14),
-        legend.title = element_blank()) 
+        legend.title = element_blank(),
+        plot.tag.position = "bottomleft") 
 ```
@@
   theme_minimal() +
-  labs(title = "Rate of Monkeypox Infections", x = "Month (2022)", 
+  labs(title = "Rate of Monkeypox Infections", 
+       tag = "Figure 2",
+       x = "Month (2022)", 
        y = "Monkeypox Rate (per 100k)") +

### 2022-11-21 demog_example_mpx.Rmd

---
title: "MPX Example"
output: html_notebook
---

## Ecological analysis using census data

This is an example of creating a metric from the census data relating to percent of population within a country/region residing in small cities

You'll probably need to change the join datasets to match the name and the join by columns that you have set up.  But hopefully this enough to give you an idea.

```{r recode_metric, echo=TRUE, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(magrittr)
library(stringr)
library(tidyr)
library(purrr)
library(lubridate)
library(plotly)

euro_census <- read_csv("https://raw.githubusercontent.com/PHW290/phw251_projectdata/main/euro_census_stats_mod.csv") %>%
  rename_all(. %>% tolower() %>% gsub(" |-|/", "_", .))

mpx_cases <- read_csv("https://raw.githubusercontent.com/PHW290/phw251_projectdata/main/euro_mpx_cases_mod.csv") %>%
  rename_all(. %>% tolower() %>% gsub(" |-|/", "_", .)) %>%
  mutate(month = floor_date(daterep, "month")) %>%
  select(-source, daterep, countryexp, country_code = countrycode, confcases, month)

# collapsing the res_pop into three categories, determining the % of pop in each category for each country
# then categorizing each country into a 3-level metric

census_metric <- euro_census%>%
  mutate(res_size_cat = case_when(res_pop %in% c("0-1000", "1000-9999", "10000-99999") ~ "small",
                                  res_pop %in% c("100000-499999") ~ "medium",
                                  res_pop %in% c("500000-999999", "GE1000000") ~ "large",
                                  TRUE ~ "unknown")
         ) %>%
  group_by(country_code, res_size_cat) %>%
  summarize(group_pop = sum(pop, na.rm = T)) %>%
  mutate(country_pop = sum(group_pop, na.rm = T),
         group_pct = group_pop / country_pop) %>%
  ungroup() %>%
  filter(res_size_cat == "small") %>%
  mutate(pct_small = case_when(group_pct >= 0.80 ~ "Greater than 80% in Small Cities",
                                group_pct >= 0.65 ~ "Between 65% and 80% in Small Cities",
                                group_pct < 0.65 ~ "Less than 65% in Small Cities"))

# merging the demographic metric and population data set with case counts and calculating rates by cat and month

res_size <- mpx_cases %>%
  group_by(country_code, month) %>%
  summarize(cases_by_month = sum(confcases, na.rm = T)) %>%
  left_join(., census_metric, by="country_code") %>%
  group_by(pct_small, month) %>%
  summarize(cases_by_monthcat = sum(cases_by_month, na.rm = T),
            cat_pop = max(country_pop, na.rm = T)) %>%
  mutate(size_cat_rate = (cases_by_monthcat / cat_pop) * 1000000) %>%
  ungroup()

fig <- plot_ly(res_size, x = ~month, y = ~size_cat_rate, type = 'scatter', mode = 'lines', linetype = ~pct_small) %>%
  layout(
    title = 'MPX Case Rates in Europe by Percent of Population Living in Small Cities, May-August 2022',
    xaxis = list(title = '', dtick = "M1", tickformat="%b<br>%Y"),
    yaxis = list(title = 'Case Rates per 1M population'),
    legend=list(orientation = 'h')
  )

fig
```

### 2022-11-21 MS4.Rmd
Adding figure labels

@@
 ### Bar Graph
-In the bar graph above we see the breakdown of monthly Monkeypox infections in Europe between May and August of this year.
+In the bar graph (fig.
+1) above we see the breakdown of monthly Monkeypox infections in Europe between May and August of this year.
 The majority of cases occured in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
@@
 ### Line Plot
-As a follow up to the bar graph above, we wanted to look at the rate of Monkeypox infections per region.
-In the bar graph above, Western Europe had the greatest number of Monkeypox infections however when we look at the rate of infections, we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.
+As a follow up to the bar graph (fig.
+1) above, we wanted to look at the rate of Monkeypox infections per region.
+In the bar graph above, Western Europe had the greatest number of Monkeypox infections however when we look at the rate of infections (fig.
+2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.

### 2022-11-21 MS4.Rmd
Table 1.2 addition

--- 2022-11-17 MS4.Rmd
@@
 library(kableExtra)
+library(janitor)
+
 ```
@@
 
+### Table 1.2
+```{r Table Visualization}
+table_1.2df <- euro_mpx_cases %>% select( EU_region, "0-1000", "1000-9999",
+                                          "10000-99999", "100000-199999",
+                                          "200000-499999" , "500000-999999",
+                                          "GE1000000")
+table_1.2df1<- pivot_longer(table_1.2df, cols =  3:9, names_to = "city_pop",
+                            values_to = "counts" )
+
+#I believe the issue is with the formation of the information below
+table_1.2df2 <- table_1.2df1 %>% 
+  group_by(EU_region, city_pop) %>% 
+  summarize(count = sum(counts))
+
+table_1.2 <- table_1.2df1 %>%
+  tabyl(city_pop, EU_region)
+
+```
+
 ```{r}
@@
   theme_minimal() +
-  labs(title = "Monthly Cases of Monkeypox in Europe", x = "Month (2022)", 
+  labs(title = "Figure 1: Monthly Cases of Monkeypox in Europe", x = "Month (2022)", 
        y = "Monkeypox Cases") +
@@
 
-### Bar Graph
+### Figure 1
 In the bar graph above we see the breakdown of monthly Monkeypox infections in Europe between May and August of this year.
-The majority of cases occured in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
+The majority of cases occurred in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
 Over the course of 4 months, the total number of Monkeypox cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214
@@
   theme_minimal() +
-  labs(title = "Rate of Monkeypox Infections", x = "Month (2022)", 
+  labs(title = "Figure 2: Rate of Monkeypox Infections", x = "Month (2022)", 
        y = "Monkeypox Rate (per 100k)") +
@@
 
-### Line Plot
-As a follow up to the bar graph above, we wanted to look at the rate of Monkeypox infections per region.
+### Figure 2
+Following the graph from Figure 1, we wanted to look at the rate of Monkeypox infections per region.
 In the bar graph above, Western Europe had the greatest number of Monkeypox infections however when we look at the rate of infections, we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.

### 2022-11-21 MS4.Rmd

--- 2022-11-21 MS4.Rmd
@@
 ```{r}
+table_1.2df <- euro_mpx_cases %>% select( EU_region, "0-1000", "1000-9999", 
+                                          "10000-99999", "100000-199999", 
+                                          "200000-499999" , "500000-999999", 
+                                          "GE1000000") 
+table_1.2df1<- pivot_longer(table_1.2df, cols = 3:9, names_to = "city_pop", 
+                            values_to = "counts" )
+
+#I believe the issue is with the formation of the information below table_1.2df2 <- table_1.2df1 %>% group_by(EU_region, city_pop) %>% summarize(count = sum(counts))
+
+table_1.2 <- table_1.2df1 %>% tabyl(city_pop, EU_region)
+
+table_1.2
+```
+
+### Table 1.2
+```{r}
 euro_mpx_cases %>% 
@@
   mutate(Total_euro_pop = sum(country_pop_2022)) 
-
 graph1 <- euro_mpx_cases %>%
@@
   mutate("monthly_rate_region" = round((region_monthly_cases/Total_region_pop * 100000),2))
-
 ggplot(data=graph1, aes(x=floor_month, y=monthly_rate_region, group= EU_region)) +

### 2022-11-21 MS4.Rmd
Table with gender and age

--- 2022-11-21 MS4.Rmd
@@
 library(kableExtra)
+library(janitor)
+library(DT)
 ```
@@
 2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.
+
+\newpage
+
+```{r}
+euro_census_stats2 <- read_csv("files/euro_census_stats.csv")
+euro_census_stats2 <- euro_census_stats2 %>%
+  mutate(AGE = case_when(AGE %in% c("Y_LT15") ~ "Age < 15",
+                                  AGE %in% c("Y15-29") ~ "Age 15-29",
+                                  AGE %in% c("Y30-49") ~ "Age 30-49",
+                                  AGE %in% c("Y50-64") ~ "Age 50-64",
+                                  AGE %in% c("Y65-84") ~ "Age 65-84",
+                                  AGE %in% c("Y_GE85") ~ "Age 85 and Older")) %>%
+  filter(pop > 0) %>%
+  select(COUNTRY_CODE, SEX, AGE, pop) %>%
+  ungroup() %>%
+  group_by(COUNTRY_CODE, SEX, AGE) %>%
+  summarise("age_category_count" = sum(pop)) %>%
+  ungroup() %>%
+  group_by(COUNTRY_CODE) %>%
+  mutate(total_pop = sum(age_category_count)) %>% ungroup()
+
+euro_mpx_cases2 <- read_csv("files/euro_mpx_cases.csv")
+euro_mpx_cases2 <- euro_mpx_cases2 %>%
+  rename(COUNTRY_CODE = CountryCode) %>%
+  group_by(COUNTRY_CODE) %>% 
+  summarize(total_cases = sum(ConfCases))
+
+demo_data <- left_join(euro_census_stats2, euro_mpx_cases2, by = "COUNTRY_CODE") %>%
+  drop_na(total_cases) %>%
+   mutate(COUNTRY_CODE = case_when(COUNTRY_CODE %in% c("AT") ~ "Austria",
+                                   COUNTRY_CODE %in% c("BE") ~ "Belgium",
+                                   COUNTRY_CODE %in% c("BG") ~ "Bulgaria",
+                                   COUNTRY_CODE %in% c("HR") ~ "Croatia",
+                                   COUNTRY_CODE %in% c("CY") ~ "Cyprus",
+                                   COUNTRY_CODE %in% c("CZ") ~ "Czechia",
+                                   COUNTRY_CODE %in% c("EE") ~ "Estonia",
+                                   COUNTRY_CODE %in% c("FI") ~ "Finland",
+                                   COUNTRY_CODE %in% c("FR") ~ "France",
+                                   COUNTRY_CODE %in% c("EL") ~ "Greece",
+                                   COUNTRY_CODE %in% c("HU") ~ "Hungary",
+                                   COUNTRY_CODE %in% c("IE") ~ "Ireland",
+                                   COUNTRY_CODE %in% c("LV") ~ "Latvia",
+                                   COUNTRY_CODE %in% c("LU") ~ "Luxembourg",
+                                   COUNTRY_CODE %in% c("MT") ~ "Malta",
+                                   COUNTRY_CODE %in% c("PL") ~ "Poland",
+                                   COUNTRY_CODE %in% c("PT") ~ "Portugal",
+                                   COUNTRY_CODE %in% c("RO") ~ "Romania",
+                                   COUNTRY_CODE %in% c("SK") ~ "Slovakia",
+                                   COUNTRY_CODE %in% c("SI") ~ "Slovenia")) %>%
+  group_by(COUNTRY_CODE, SEX, AGE, total_cases, total_pop) %>%
+  mutate('age_cat_pct' = round((age_category_count/total_pop*100),digits = 2)) %>%
+  select(COUNTRY_CODE, SEX, AGE, total_cases, total_pop, age_cat_pct) %>%
+  pivot_wider(names_from = AGE, values_from = age_cat_pct) %>%
+  group_by(COUNTRY_CODE, SEX) %>%
+  select(-c(total_pop)) %>%
+  arrange(desc(total_cases))
+
+kable(demo_data, "latex", booktabs = T,
+      col.names = c("Country",
+                    "Sex",
+                    "MPX Cases",
+                    "Age < 15 (%)",
+                    "Age 15-29 (%)",
+                    "Age 30-49 (%)",
+                    "Age 50-64 (%)",
+                    "Age 65-84 (%)",
+                    "Age 85 and Older (%)"),
+      caption = "European Countries by Sex and Age Category") %>%
+  kable_styling(latex_options = "HOLD_position")
+
+```

### 2022-11-22 MS4.Rmd
Final Submittal for MS4

--- 2022-11-21 MS4.Rmd
@@
 library(janitor)
-
+library(DT)
 ```
@@
 
-```{r}
-table_1.2df <- euro_mpx_cases %>% select( EU_region, "0-1000", "1000-9999", 
-                                          "10000-99999", "100000-199999", 
-                                          "200000-499999" , "500000-999999", 
-                                          "GE1000000") 
-table_1.2df1<- pivot_longer(table_1.2df, cols = 3:9, names_to = "city_pop", 
-                            values_to = "counts" )
+```{r Population Density Demographic Table}
+euro_census_stats1 <- read_csv("files/euro_census_stats.csv")
 
-#I believe the issue is with the formation of the information below table_1.2df2 <- table_1.2df1 %>% group_by(EU_region, city_pop) %>% summarize(count = sum(counts))
+euro_census_stats2 <- euro_census_stats1 %>%
+  group_by(COUNTRY_CODE, RES_POP) %>%
+  summarise(census_pop = sum(pop)) %>%
+  rename(CountryCode = COUNTRY_CODE) %>% 
+  mutate(EU_region = case_when(
+    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
+    CountryCode %in% sr_northern_e ~ "Northern Europe",
+    CountryCode %in% sr_southern_e ~ "Southern Europe",
+    CountryCode %in% sr_western_e ~ "Western Europe",
+    CountryCode %in% sr_western_a ~ "Western Asia"
+    ))
 
-table_1.2 <- table_1.2df1 %>% tabyl(city_pop, EU_region)
+euro_census_stats2a <- euro_census_stats2 %>%
+  group_by(EU_region, RES_POP) %>%
+  summarize(pop_density_count = sum(census_pop)) %>%
+  na.omit()
 
-table_1.2
+Table_2a <- pivot_wider(euro_census_stats2a,
+            names_from = EU_region,
+            values_from = pop_density_count)
+
+kable(Table_2a,
+      col.names = c("Population Density", "Eastern Europe", "Northern Europe",
+                    "Southern Europe","Western Asia","Western Europe"),
+      caption = "Population Demographic (Population Density) per Region") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+## Table 4
+
+Table 4 illustrates the strata of individuals living within each region of the EU by the population density of that region.
+
+\newpage
+
+```{r Age Demographic Table}
+euro_census_stats2b <- euro_census_stats1 %>%
+  group_by(COUNTRY_CODE, AGE) %>%
+  summarise(census_pop = sum(pop)) %>%
+  rename(CountryCode = COUNTRY_CODE) %>% 
+  mutate(EU_region = case_when(
+    CountryCode %in% sr_eastern_e ~ "Eastern Europe",
+    CountryCode %in% sr_northern_e ~ "Northern Europe",
+    CountryCode %in% sr_southern_e ~ "Southern Europe",
+    CountryCode %in% sr_western_e ~ "Western Europe",
+    CountryCode %in% sr_western_a ~ "Western Asia"
+    ))
+
+euro_census_stats2b <- euro_census_stats2b %>%
+  group_by(EU_region, AGE) %>%
+  summarize(pop_density_count = sum(census_pop)) %>%
+  na.omit()
+
+Table_2b <- pivot_wider(euro_census_stats2b,
+            names_from = EU_region,
+            values_from = pop_density_count)
+
+kable(Table_2b,
+      col.names = c("Age Group", "Eastern Europe", "Northern Europe",
+                    "Southern Europe","Western Asia","Western Europe"),
+      caption = "Age Groups (Age Density) per Region") %>%
+  kable_styling(latex_options = "HOLD_position")
 ```
 
-### Table 1.2
+## Table 5
+
+Table 5 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
+
+\newpage
+
 ```{r}
@@
 In the bar graph (fig.
-1) above we see the breakdown of monthly Monkeypox infections in Europe between May and August of this year.
-The majority of cases occured in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
-Over the course of 4 months, the total number of Monkeypox cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214
+1) above we see the breakdown of monthly MPX infections in Europe between May and August of this year.
+The majority of cases occurred in Western Europe and Southern Europe as evidenced by the respectively colored purple and green regions in the bar graph.
+Over the course of 4 months, the total number of MPX cases by region were Western Asia = 4, Eastern Europe = 277, Northern Europe = 619, Southern Europe = 7958, and Western Europe = 8214
 
@@
 As a follow up to the bar graph (fig.
-1) above, we wanted to look at the rate of Monkeypox infections per region.
-In the bar graph above, Western Europe had the greatest number of Monkeypox infections however when we look at the rate of infections (fig.
-2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 monkeypox cases per 100,000 people in the month of July.
+1) above, we wanted to look at the rate of MPX infections per region.
+In the bar graph above, Western Europe had the greatest number of MPX infections however when we look at the rate of infections (fig.
+2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 MPX cases per 100,000 people in the month of July.
+
+\newpage
+
+```{r}
+euro_census_stats2 <- read_csv("files/euro_census_stats.csv")
+euro_census_stats2 <- euro_census_stats2 %>%
+  mutate(AGE = case_when(AGE %in% c("Y_LT15") ~ "Age < 15",
+                                  AGE %in% c("Y15-29") ~ "Age 15-29",
+                                  AGE %in% c("Y30-49") ~ "Age 30-49",
+                                  AGE %in% c("Y50-64") ~ "Age 50-64",
+                                  AGE %in% c("Y65-84") ~ "Age 65-84",
+                                  AGE %in% c("Y_GE85") ~ "Age 85 and Older")) %>%
+  filter(pop > 0) %>%
+  select(COUNTRY_CODE, SEX, AGE, pop) %>%
+  ungroup() %>%
+  group_by(COUNTRY_CODE, SEX, AGE) %>%
+  summarise("age_category_count" = sum(pop)) %>%
+  ungroup() %>%
+  group_by(COUNTRY_CODE) %>%
+  mutate(total_pop = sum(age_category_count)) %>% ungroup()
+
+euro_mpx_cases2 <- read_csv("files/euro_mpx_cases.csv")
+euro_mpx_cases2 <- euro_mpx_cases2 %>%
+  rename(COUNTRY_CODE = CountryCode) %>%
+  group_by(COUNTRY_CODE) %>% 
+  summarize(total_cases = sum(ConfCases))
+
+demo_data <- left_join(euro_census_stats2, euro_mpx_cases2, by = "COUNTRY_CODE") %>%
+  drop_na(total_cases) %>%
+   mutate(COUNTRY_CODE = case_when(COUNTRY_CODE %in% c("AT") ~ "Austria",
+                                   COUNTRY_CODE %in% c("BE") ~ "Belgium",
+                                   COUNTRY_CODE %in% c("BG") ~ "Bulgaria",
+                                   COUNTRY_CODE %in% c("HR") ~ "Croatia",
+                                   COUNTRY_CODE %in% c("CY") ~ "Cyprus",
+                                   COUNTRY_CODE %in% c("CZ") ~ "Czechia",
+                                   COUNTRY_CODE %in% c("EE") ~ "Estonia",
+                                   COUNTRY_CODE %in% c("FI") ~ "Finland",
+                                   COUNTRY_CODE %in% c("FR") ~ "France",
+                                   COUNTRY_CODE %in% c("EL") ~ "Greece",
+                                   COUNTRY_CODE %in% c("HU") ~ "Hungary",
+                                   COUNTRY_CODE %in% c("IE") ~ "Ireland",
+                                   COUNTRY_CODE %in% c("LV") ~ "Latvia",
+                                   COUNTRY_CODE %in% c("LU") ~ "Luxembourg",
+                                   COUNTRY_CODE %in% c("MT") ~ "Malta",
+                                   COUNTRY_CODE %in% c("PL") ~ "Poland",
+                                   COUNTRY_CODE %in% c("PT") ~ "Portugal",
+                                   COUNTRY_CODE %in% c("RO") ~ "Romania",
+                                   COUNTRY_CODE %in% c("SK") ~ "Slovakia",
+                                   COUNTRY_CODE %in% c("SI") ~ "Slovenia")) %>%
+  group_by(COUNTRY_CODE, SEX, AGE, total_cases, total_pop) %>%
+  filter(total_cases > 100) %>%
+  mutate('age_cat_pct' = round((age_category_count/total_pop*100),digits = 2)) %>%
+  select(COUNTRY_CODE, SEX, AGE, total_cases, total_pop, age_cat_pct) %>%
+  group_by(COUNTRY_CODE, AGE, total_cases) %>%
+  summarise(age_cat_pct = sum(age_cat_pct)) %>%
+  arrange(desc(AGE))
+
+demo_data %>% 
+  ggplot(., aes(x = COUNTRY_CODE, y = age_cat_pct)) +
+  geom_bar(aes(fill = AGE), stat = "identity") +
+  theme_minimal() +
+  labs(title = "Age Distribution Among Countries with Monkeypox Cases",
+       subtitle = "Countries with at least 100 confirmed cases",
+       tag = "Figure 3",
+       x = "Country", 
+       y = "Percentage") +
+  theme(axis.title = element_text(size = 10),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(size = 14),
+        legend.title = element_blank(),
+        plot.tag.position = "bottomleft")
+
+```
+
+### Stacked Bargraph Age Distribution
+
+Figure 3 includes countries that experience at least 100 cases of MPX.
+The figure parameters include stacked bar graphs that visually represent the age brackets from which a positive case is observed.
+The 30-49 age range is observed to have more positive cases across each country than any of the other age brackets.
+Given the consistency within this observation, the investigators may find that countries with populations with thriving numbers of 30-49-year-old's are affected more by MPX than regions that do not.
+Moreover, the countries listed in this figure are predominantly within the Western Region of Europe; Western Europe includes six countries.

### 2022-11-28 MS5.Rmd
Working on MS5

--- 2022-11-22 MS4.Rmd
@@
 ---
-title: "Scenario 2 - MPX in the EU"
+title: "Final Report: Scenario 2 - MPX in the EU"
 author: "Dawson Coblin and Jose Lazo"
-date: '2022-11-21'
+date: '2022-11-28'
 output: pdf_document
@@
 
-# Monkeypox (MPX) Infection Rates in the European Union
+# Problem Statement
 
-This report aims to prepare data to analyze and assess the impact of MPV
-within\
-each country of the EU, and understand how case rates may differ by
-region and other various demographic factors. These findings will aid
-the implementation and appropriate response to the epidemic to local
-health departments.
+As the world begins to recover from the effects of COVID, the threat of the monkeypox virus (MPX) has public health officials preparing for the worst.
+Throughout Europe, cases of MPX have begun to rise and it is imperative to understand the groups at highest risk.
+This report aims to prepare data to analyze and assess the impact of MPX
+within each country of the EU, and understand how case rates may differ by region and other various demographic factors.
+These findings will aid the implementation and appropriate response to the epidemic to local
+health departments. Our team hypothesizes that monkeypox will be prevalent among the very young and old as is common for many infectious diseases.
 
-### Legend of Tasks and Milestones
+# Methods
 
-| Milestone   | Description of Section                                           |
-|-----------------------|-------------------------------------------------|
-| Milestone 1 | Group Agreement                                                  |
-| Milestone 2 | Data-set Imported into R; Identify Key Data Elements and Types   |
-| Milestone 3 | Clean Data-set Create Descriptive Statistics and Data Dictionary |
-| Milestone 4 | Data Visualization                                               |
+Datasets with information on confirmed cases of monkeypox per country as well as census data from Europe were used to prepare this report.
+Countries were grouped by region based on world_country_region data set to describe the cases per region.
 
-Milestone 3 Sub-sections:\
-1.Sub-setting Rows and Columns as Needed- Clean individual Data Frames\
-2.Summarizing Data and Creating New Vectors, Organize Vectors for Use\
-3.Clean Variables and Combined Data-sets\
-4.Descriptive Stats of EU Data and Data Dictionary
-
-\newpage
-
-### 1) Milestone 1: Group Agreement
-
-Group R: Team Logistics Agreement
-
-9/19/22
-
-1.  Project option selection
-
-Scenario 2: Monkeypox (MPX) infection rates in the European Union
-
-2.  Create git repository
-
-<https://github.com/dcoblin/251.git>
-
-3.  Roles and responsibilities
-
-```{=html}
-<!-- -->
-```
-a.  Team's preferred communication method
-
-WhatsApp
-
-b.  Team's preferred meeting times and frequency
-
-Weekends/Post 6 PM on weekdays
-
-c.  Team's preferred method for tracking progress
-
-Google docs
-
-d.  Point person for contacting course facilitators with questions
-
-Jose/Dawson
-
-e.  Point person for submitting all milestones
-
-Dawson/Jose
-
-\newpage
-
-### 2) Milestone 2: Data-set Imported into R; Identify Key Data Elements and Types
+Within the euro_mpx_cases dataset, relevant fields include country, country code, and confirmed cases. A curated demographic dataset included information on age, sex, and number of people per strata.
 
@@
 
-```{r}
+```{r, include=FALSE}
 euro_mpx_cases <- read_csv("files/euro_mpx_cases.csv")
@@
 
-\newpage
-
-## Milestone 3: 1. Sub-setting Rows and Columns as Needed- Clean individual Data Frames
-
-### Data Frame Cleaning
-
-```{r}
+```{r, include=FALSE}
 #Cleaning the cases df
@@
                                                %in% euro_country_vector, ]
-
-head(euro_pop_denominators,3)
 
@@
 
-head(euro_census_stats, 3)
 ```
 
-\newpage
-
-## Milestone 3: 2. Summarizing Data and Creating New Vectors, Organize Vectors for Use
-
-```{r}
+```{r, include=FALSE}
 #We then assign counties with a region code in a new column a list "EU_region,"
@@
     ))
-
-head(euro_mpx_cases,3)
 ```
 
-\newpage
-
-## Milestone 3: 3. Clean Variables and Combined Data-sets
-
-```{r}
+```{r, include=FALSE}
 
@@
 
-head(euro_mpx_cases,3)
 ```
 
-\newpage
-
-## Milestone 3: 4. Descriptive Stats of EU Data and Data Dictionary
-
-### Descriptive stats:
-
-We find that the census information for 7 of the 29 countries in the EU
-are not available. These "0" values were replaced with "NA" to prevent
-graphing issues.
-
-### Region Data:
-
-We see that the regions in EU include various numbers of countries, thus
-having unequal populations. A useful metric to create a graph will be
-the calculation of cases per 100,000. This will allow us to use 1
-categorical variable (EU Region) to a quantitative variable (number of
-confirmed cases)
-
-We compiled the relevant confirmed case information per region in the
-EU.
-
-### The pertinent information per EU region is listed below:
+# Results
 
@@
 
-## Data Dictionary
+In table 1, we see that the regions in EU include various numbers of countries, thus
+having unequal populations. A useful metric to create a graph will be
+the calculation of cases per 100,000. This will allow us to use 1
+categorical variable (EU Region) to a quantitative variable (number of
+confirmed cases). We compiled the relevant confirmed case information per region in the
+EU.
 
-```{r, echo=FALSE}
-#Create the Data Dictionary
-
-names <- names(euro_mpx_cases)
-data_type <- sapply(euro_mpx_cases,class)
-description <- c("Reporting time period with the floor month", #floor_month
-                 "Abbreviation of Country", # CountryCode
-                 "The current population of the country for 2022", # country_pop_2022
-                 "Total confirmed MPX cases grouped by month", # cases_grouped_monthly
-                 "The designated region the country is located", # EU_region
-                 "Sum of the four months per country code", # sum_region_cases
-                 "The Sum of population 0-1000", # 0-1000
-                 "The Sum of population 1000-9999", # 1000-9999
-                 "The Sum of population 10000-99999", # 10000-99999
-                 "The Sum of population 100000-199999", # 100000-199999
-                 "The Sum of population 200000-499999", # 200000-499999
-                 "The Sum of population 500000-999999", # 500000-999999
-                 "The Sum of population GE1000000" # GE1000000
-)
-
-data_dic <- as.data.frame(cbind(names, data_type, description)) %>% 
-  select(-names)
-
-kable(data_dic,
-      col.names = c("Data Type", 
-                    "Description"),
-      caption = "Data Dictionary"
-      ) %>%
-  kable_styling(latex_options = "HOLD_position")
-```
-
-\newpage
-
-### Milestone 4: Visualizations
-
-```{r Population Density Demographic Table}
+```{r Population Density Demographic Setup, include=FALSE}
 euro_census_stats1 <- read_csv("files/euro_census_stats.csv")
@@
             values_from = pop_density_count)
+```
 
+```{r Population Density Demographic Table, echo=FALSE}
 kable(Table_2a,
@@
 ```
-## Table 4
 
-Table 4 illustrates the strata of individuals living within each region of the EU by the population density of that region.
+Table 2 illustrates the strata of individuals living within each region of the EU by the population density of that region.
 
-\newpage
-
-```{r Age Demographic Table}
+```{r Age Demographic Setup, include=FALSE}
 euro_census_stats2b <- euro_census_stats1 %>%
@@
             values_from = pop_density_count)
+```
+
+```{r Age Demographic table, echo=FALSE}
 
@@
 
-## Table 5
+Table 3 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
 
-Table 5 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
-
-\newpage
-
-```{r}
+```{r, echo=FALSE}
 euro_mpx_cases %>% 
@@
 
-### Bar Graph
 In the bar graph (fig.
@@
 
-\newpage
-
-```{r}
+```{r, echo=FALSE}
 euro_mpx_cases <- inner_join(euro_mpx_cases, Table_1.1, by= "EU_region") %>%
@@
 
-### Line Plot
 As a follow up to the bar graph (fig.
@@
 
-\newpage
-
-```{r}
+```{r, include=FALSE}
 euro_census_stats2 <- read_csv("files/euro_census_stats.csv")
@@
   arrange(desc(AGE))
+```
+
+```{r, echo=FALSE}
 
@@
 
-### Stacked Bargraph Age Distribution
-
 Figure 3 includes countries that experience at least 100 cases of MPX.
@@
 Moreover, the countries listed in this figure are predominantly within the Western Region of Europe; Western Europe includes six countries.
+
+# Discussion

### 2022-11-28 MS5.Rmd
Cleaned the methods section and inserted the discussion section.

@@
 
-As the world begins to recover from the effects of COVID, the threat of the monkeypox virus (MPX) has public health officials preparing for the worst.
-Throughout Europe, cases of MPX have begun to rise and it is imperative to understand the groups at highest risk.
-This report aims to prepare data to analyze and assess the impact of MPX
-within each country of the EU, and understand how case rates may differ by region and other various demographic factors.
-These findings will aid the implementation and appropriate response to the epidemic to local
-health departments. Our team hypothesizes that monkeypox will be prevalent among the very young and old as is common for many infectious diseases.
+As the world begins to recover from the effects of COVID, the threat of the so-called monkeypox virus (MPX) has public health officials preparing for the worst.
+Throughout Europe (EU), cases of MPX have begun to rise, and it is imperative to understand the groups at the highest risk for infection.
+This report aims to prepare data to analyze and assess the impact of MPX within each country of the EU, understand how case rates may differ by region, and find potential links to various demographic factors.
+These findings will aid local health departments' implementation and appropriate response to the epidemic.
+Our team hypothesizes that MPX will be prevalent among young (under 30 years old) and older populations (over 64 years old), as is typical for many infectious diseases.
 
@@
 
-Datasets with information on confirmed cases of monkeypox per country as well as census data from Europe were used to prepare this report.
-Countries were grouped by region based on world_country_region data set to describe the cases per region.
-
-Within the euro_mpx_cases dataset, relevant fields include country, country code, and confirmed cases. A curated demographic dataset included information on age, sex, and number of people per strata.
+To prepare this report, use data sets for the EU, explicitly including information on confirmed cases of MPX per country (euro_mpx_cases.csv) and census data (euro_census_stats.csv).
+The trajectory of this investigation incorporated four sets of data manipulation:
+(1) Grouping countries into regions;
+(2) Assessing necessary census information;
+(3) Compiling data sets with confirmed case demographics;
+(4) Creating tables and figures for visual analysis.
+Countries were grouped by region based on a region's data set (world_country_regions.csv) to enhance the representation of positive cases per region.
+Within the euro_mpx_cases data set, relevant fields include country, country code, and confirmed cases.
+A curated demographic data set included information on age, sex, and the number of people per strata.
+As an aside with the census data, we chose to initially assess any "missing" data as "NA" when paired with a country.
+However then decided to print information with omit-NA's to allow aggregate fields to be added together for "Table 2: Population Demographic (Population Density) per Region."
 
@@
 
-In table 1, we see that the regions in EU include various numbers of countries, thus
+In Table 1, we see that the regions in EU include various numbers of countries, thus
 having unequal populations. A useful metric to create a graph will be
-the calculation of cases per 100,000. This will allow us to use 1
+the calculation of cases per 100,000. This will allow us to use one
 categorical variable (EU Region) to a quantitative variable (number of
@@
 
-In the bar graph (fig.
+In the bar graph (Fig.
 1) above we see the breakdown of monthly MPX infections in Europe between May and August of this year.
@@
 
-As a follow up to the bar graph (fig.
+As a follow up to the bar graph (Fig.
 1) above, we wanted to look at the rate of MPX infections per region.
-In the bar graph above, Western Europe had the greatest number of MPX infections however when we look at the rate of infections (fig.
-2), we see that over the course of the 4 months, Southern Europe had the greatest rate of infection with a peak of 2.84 MPX cases per 100,000 people in the month of July.
+In the bar graph above, Western Europe had the greatest number of MPX infections however when we look at the rate of infections (Fig.
+2), we see that over the course of the four months, Southern Europe had the greatest rate of infection with a peak of 2.84 MPX cases per 100,000 people in the month of July.
 
@@
 # Discussion
+
+Our analysis of Table 1 and Figure 1 reveal the impacted regions for MPX though the study period (May through August 2022) occurred in the Southern and Western Regions of Europe.
+Within the EU, these two regions comprise 13 of its 29 total countries, roughly 71.8% of the total EU population (extrapolated using Table 1).
+
+Through empirical analysis in Figure 3, we are not confident of the initial hypothesis regarding younger (under 30 years old) and older (over 64 years old) populations having a higher risk of infection.
+We observe that the age group for 30-64-year-olds (broken into two categories of 30-49 and 50-64 years) contributes over 50% of MPX cases in countries with a minimum of 100 confirmed MPX cases (six countries from the 29 countries in the EU).
+Given that the 30-64 age group appears to burden more patients than those below 29 years old and above 64 years old for all six of the countries that have 100+ MPX cases, we believe that the expansion of this finding will be evident throughout the remaining 23 countries in the EU.
+
+Given the analysis of this information, health department leadership should begin educational programs and outreach focusing on those 30-64 aged individuals.

### 2022-11-28 MS5.Rmd
Update to methods and insert of discussion section

@@
 
-To prepare this report, use data sets for the EU, explicitly including information on confirmed cases of MPX per country (euro_mpx_cases.csv) and census data (euro_census_stats.csv).
+To prepare this report our team uses data sets for the EU, explicitly including information on confirmed cases of MPX per country (euro_mpx_cases.csv) and census data (euro_census_stats.csv).
 The trajectory of this investigation incorporated four sets of data manipulation:

### 2022-12-06 Final Report: Scenario 2 - MPX in the EU.Rmd
Edited figure1,2,3

--- 2022-11-28 MS5.Rmd
@@
 ```{r, echo=FALSE}
-euro_mpx_cases %>% 
+euro_mpx_cases1.5 <- euro_mpx_cases %>%
+  select(floor_month, EU_region, cases_grouped_monthly) %>%
+  group_by(floor_month, EU_region) %>%
+  summarize(cases_grouped_monthly = sum(cases_grouped_monthly))
+
+euro_mpx_cases1.5 %>%
   ggplot(., aes(x = floor_month, y = cases_grouped_monthly)) +
@@
   geom_point() +
-  geom_label(label= graph1$monthly_rate_region, nudge_y = 0.1) +
+  geom_text(label= graph1$monthly_rate_region, nudge_y = 0.1, check_overlap = T) +
   theme_minimal() +
@@
   geom_bar(aes(fill = AGE), stat = "identity") +
+  geom_text(aes(label= age_cat_pct), position=position_stack(vjust=0.5), size=2.5) +
   theme_minimal() +

### 2022-12-07 MS5.Rmd
TEST

@@
 author: "Dawson Coblin and Jose Lazo"
-date: '2022-11-28'
+date: '2022-12-11'
 output: pdf_document
@@
 Our team hypothesizes that MPX will be prevalent among young (under 30 years old) and older populations (over 64 years old), as is typical for many infectious diseases.
+We believe that countries with larger proportions of young and old populations will have more cases of MPX than those regions that have countries with higher middle aged population proportions.
 
@@
 (4) Creating tables and figures for visual analysis.
+
 Countries were grouped by region based on a region's data set (world_country_regions.csv) to enhance the representation of positive cases per region.
@@
 As an aside with the census data, we chose to initially assess any "missing" data as "NA" when paired with a country.
-However then decided to print information with omit-NA's to allow aggregate fields to be added together for "Table 2: Population Demographic (Population Density) per Region."
+However then decided to print information with omit-NA's to allow aggregate fields to be added together for review of population demographic (Population Density) per Region.
 
@@
 
+\newpage
+
 # Results
@@
 Countries_per_Region = c(6, 9, 7, 1, 6)
-Table_1.1 <- as.data.frame(cbind(Table_1.1, Countries_per_Region))
+Cumulative_case_rate_per_100k = c(.31, 1.65, 5.94, .44, 4.29)
+Table_1.1 <- as.data.frame(cbind(Table_1.1, Countries_per_Region , 
+                                 Cumulative_case_rate_per_100k))
 ```
@@
                     "Total Region Population",
-                    "Countries Per Region"),
-      caption = "Monkeypox Cases Per Region"
+                    "Countries Per Region",
+                    "Case Rate per 100k"),
+      caption = "MPX Cases Per Region"
       ) %>%
@@
 confirmed cases). We compiled the relevant confirmed case information per region in the
-EU.
+EU with the cumulative case rate per 100k represented in the 3rd column.
+
+\newpage
 
@@
 
-```{r Population Density Demographic Table, echo=FALSE}
+```{r Population Density Demographic Table, eval=FALSE, include=FALSE}
 kable(Table_2a,
@@
   kable_styling(latex_options = "HOLD_position")
+
+#Table 2 illustrates the strata of individuals living within each region of the EU by the population density of that region.
 ```
-
-Table 2 illustrates the strata of individuals living within each region of the EU by the population density of that region.
 
@@
 
-```{r Age Demographic table, echo=FALSE}
+```{r Age Demographic table, eval=FALSE, include=FALSE}
 
@@
   kable_styling(latex_options = "HOLD_position")
+
+#Table 3 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
 ```
-
-Table 3 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
 
@@
 
-Through empirical analysis in Figure 3, we are not confident of the initial hypothesis regarding younger (under 30 years old) and older (over 64 years old) populations having a higher risk of infection.
-We observe that the age group for 30-64-year-olds (broken into two categories of 30-49 and 50-64 years) contributes over 50% of MPX cases in countries with a minimum of 100 confirmed MPX cases (six countries from the 29 countries in the EU).
-Given that the 30-64 age group appears to burden more patients than those below 29 years old and above 64 years old for all six of the countries that have 100+ MPX cases, we believe that the expansion of this finding will be evident throughout the remaining 23 countries in the EU.
+Through analysis of Figure 3, we are not confident of the initial hypothesis regarding younger (under 30 years old) and older (over 64 years old) populations proportions of country demographics causing a higher risk of infection.
+Review of the demographics associated to Figure 3 provide some incite as to the population age bracket proportions for those countries listed (produced over 100 confirmed cases).
+We observe that the age group for 30-64-year-olds (broken into two categories of 30-49 and 50-64 years) attributes to over 50 % of the population proportion for all six countries.
+When we reviewed the demographic characteristics associated to sex of the populations within each region, we felt that the use of age within our figure provided a better story into potential factors that are causing the occurrence of cases of MPX in Europe.
 

### 2022-12-07 Final Report: Scenario 2 - MPX in the EU.Rmd
Combining Final Reports

@@
 As an aside with the census data, we chose to initially assess any "missing" data as "NA" when paired with a country.
-However then decided to print information with omit-NA's to allow aggregate fields to be added together for "Table 2: Population Demographic (Population Density) per Region."
+However then decided to print information with omit-NA's to allow aggregate fields to be added together for review of population demographic (Population Density) per Region"
 
@@
 ```
+\newpage
 
@@
 confirmed cases). We compiled the relevant confirmed case information per region in the
-EU.
+EU with the cumulative case rate per 100k represented in the 3rd column.
 
@@
 ```
-
-Table 2 illustrates the strata of individuals living within each region of the EU by the population density of that region.
 
@@
 ```
-
-Table 3 summarizes the region demographics by age using data that provides information by region based on age in specific brackets.
 

### 2022-12-11 Final Report: Scenario 2 - MPX in the EU.Rmd
Further analysis of figure 3 and conclusions

@@
 
-As the world begins to recover from the effects of COVID, the threat of the so-called monkeypox virus (MPX) has public health officials preparing for the worst.
+As the world begins to recover from the effects of COVID, the threat of the monkeypox virus (MPX) has public health officials preparing for the worst.
 Throughout Europe (EU), cases of MPX have begun to rise, and it is imperative to understand the groups at the highest risk for infection.
@@
 These findings will aid local health departments' implementation and appropriate response to the epidemic.
-Our team hypothesizes that MPX will be prevalent among young (under 30 years old) and older populations (over 64 years old), as is typical for many infectious diseases.
+Our team hypothesizes that MPX will be prevalent among young (under 15 years old) and older populations (over 64 years old), as is typical for many infectious diseases.
 We believe that countries with larger proportions of young and old populations will have more cases of MPX than those regions that have countries with higher middle aged population proportions.
@@
 A curated demographic data set included information on age, sex, and the number of people per strata.
+For the ecologic analysis of monkeypox and age, countries with missing census data for individual age stratum were omitted from analysis.
 As an aside with the census data, we chose to initially assess any "missing" data as "NA" when paired with a country.
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, message=FALSE, warning=FALSE}
 #Cleaning the cases df
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, message=FALSE, warning=FALSE}
 #We then assign counties with a region code in a new column a list "EU_region,"
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 kable(Table_1.1,
@@
 confirmed cases). We compiled the relevant confirmed case information per region in the
-EU with the cumulative case rate per 100k represented in the 3rd column.
+EU with the cumulative case rate per 100k.
 
@@
 
-```{r Population Density Demographic Table, eval=FALSE, include=FALSE}
+```{r Population Density Demographic Table, eval=FALSE, include=FALSE, message=FALSE, warning=FALSE}
 kable(Table_2a,
@@
 
-```{r Age Demographic Setup, include=FALSE}
+```{r Age Demographic Setup, include=FALSE, message=FALSE, warning=FALSE}
 euro_census_stats2b <- euro_census_stats1 %>%
@@
 
-```{r Age Demographic table, eval=FALSE, include=FALSE}
+```{r Age Demographic table, eval=FALSE, include=FALSE, message=FALSE, warning=FALSE}
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 euro_mpx_cases1.5 <- euro_mpx_cases %>%
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 euro_mpx_cases <- inner_join(euro_mpx_cases, Table_1.1, by= "EU_region") %>%
@@
 
-```{r, include=FALSE}
+```{r, include=FALSE, message=FALSE, warning=FALSE}
 euro_census_stats2 <- read_csv("files/euro_census_stats.csv")
@@
                                    COUNTRY_CODE %in% c("CZ") ~ "Czechia",
+                                   COUNTRY_CODE %in% c("DK") ~ "Denmark",
                                    COUNTRY_CODE %in% c("EE") ~ "Estonia",
@@
                                    COUNTRY_CODE %in% c("FR") ~ "France",
+                                   COUNTRY_CODE %in% c("DE") ~ "Germany",
                                    COUNTRY_CODE %in% c("EL") ~ "Greece",
                                    COUNTRY_CODE %in% c("HU") ~ "Hungary",
+                                   COUNTRY_CODE %in% c("IS") ~ "Iceland",
                                    COUNTRY_CODE %in% c("IE") ~ "Ireland",
+                                   COUNTRY_CODE %in% c("IT") ~ "Italy",
                                    COUNTRY_CODE %in% c("LV") ~ "Latvia",
+                                   COUNTRY_CODE %in% c("LT") ~ "Lithuania",
                                    COUNTRY_CODE %in% c("LU") ~ "Luxembourg",
                                    COUNTRY_CODE %in% c("MT") ~ "Malta",
+                                   COUNTRY_CODE %in% c("NL") ~ "Netherlands",
+                                   COUNTRY_CODE %in% c("NO") ~ "Norway",
                                    COUNTRY_CODE %in% c("PL") ~ "Poland",
@@
                                    COUNTRY_CODE %in% c("SK") ~ "Slovakia",
-                                   COUNTRY_CODE %in% c("SI") ~ "Slovenia")) %>%
+                                   COUNTRY_CODE %in% c("SI") ~ "Slovenia",
+                                   COUNTRY_CODE %in% c("ES") ~ "Spain",
+                                   COUNTRY_CODE %in% c("SE") ~ "Sweden")) %>%
+  group_by(COUNTRY_CODE, AGE, total_cases, total_pop) %>%
+  filter(total_cases >= 100) %>%
+  mutate('age_cat_pct' = round((age_category_count/total_pop*100),digits = 2)) %>%
+  select(COUNTRY_CODE, AGE, total_cases, total_pop, age_cat_pct) %>%
+  group_by(COUNTRY_CODE, AGE, total_cases) %>%
+  summarise(age_cat_pct = sum(age_cat_pct)) %>%
+  arrange(desc(AGE))
+
+demo_data2 <- left_join(euro_census_stats2, euro_mpx_cases2, by = "COUNTRY_CODE") %>%
+  drop_na(total_cases) %>%
+  mutate(COUNTRY_CODE = case_when(COUNTRY_CODE %in% c("AT") ~ "Austria",
+                                   COUNTRY_CODE %in% c("BE") ~ "Belgium",
+                                   COUNTRY_CODE %in% c("BG") ~ "Bulgaria",
+                                   COUNTRY_CODE %in% c("HR") ~ "Croatia",
+                                   COUNTRY_CODE %in% c("CY") ~ "Cyprus",
+                                   COUNTRY_CODE %in% c("CZ") ~ "Czechia",
+                                   COUNTRY_CODE %in% c("DK") ~ "Denmark",
+                                   COUNTRY_CODE %in% c("EE") ~ "Estonia",
+                                   COUNTRY_CODE %in% c("FI") ~ "Finland",
+                                   COUNTRY_CODE %in% c("FR") ~ "France",
+                                   COUNTRY_CODE %in% c("DE") ~ "Germany",
+                                   COUNTRY_CODE %in% c("EL") ~ "Greece",
+                                   COUNTRY_CODE %in% c("HU") ~ "Hungary",
+                                   COUNTRY_CODE %in% c("IS") ~ "Iceland",
+                                   COUNTRY_CODE %in% c("IE") ~ "Ireland",
+                                   COUNTRY_CODE %in% c("IT") ~ "Italy",
+                                   COUNTRY_CODE %in% c("LV") ~ "Latvia",
+                                   COUNTRY_CODE %in% c("LT") ~ "Lithuania",
+                                   COUNTRY_CODE %in% c("LU") ~ "Luxembourg",
+                                   COUNTRY_CODE %in% c("MT") ~ "Malta",
+                                   COUNTRY_CODE %in% c("NL") ~ "Netherlands",
+                                   COUNTRY_CODE %in% c("NO") ~ "Norway",
+                                   COUNTRY_CODE %in% c("PL") ~ "Poland",
+                                   COUNTRY_CODE %in% c("PT") ~ "Portugal",
+                                   COUNTRY_CODE %in% c("RO") ~ "Romania",
+                                   COUNTRY_CODE %in% c("SK") ~ "Slovakia",
+                                   COUNTRY_CODE %in% c("SI") ~ "Slovenia",
+                                   COUNTRY_CODE %in% c("ES") ~ "Spain",
+                                   COUNTRY_CODE %in% c("SE") ~ "Sweden")) %>%
   group_by(COUNTRY_CODE, SEX, AGE, total_cases, total_pop) %>%
-  filter(total_cases > 100) %>%
+  filter(total_cases < 100) %>%
   mutate('age_cat_pct' = round((age_category_count/total_pop*100),digits = 2)) %>%
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 
-demo_data %>% 
+figure_3.1 <- demo_data %>% 
   ggplot(., aes(x = COUNTRY_CODE, y = age_cat_pct)) +
@@
        subtitle = "Countries with at least 100 confirmed cases",
-       tag = "Figure 3",
+       tag = "Figure 3A",
        x = "Country", 
@@
         plot.tag.position = "bottomleft")
+figure_3.1
 
+figure_3.2 <- demo_data2 %>% 
+  ggplot(., aes(x = COUNTRY_CODE, y = age_cat_pct)) +
+  geom_bar(aes(fill = AGE), stat = "identity") +
+  geom_text(aes(label= age_cat_pct), position=position_stack(vjust=0.5), size=1.5) +
+  theme_minimal() +
+  labs(title = "Age Distribution Among Countries with Monkeypox Cases",
+       subtitle = "Countries with less than 100 confirmed cases",
+       tag = "Figure 3B",
+       x = "Country", 
+       y = "Percentage") +
+  theme(axis.title = element_text(size = 10),
+        axis.text = element_text(size = 8),
+        axis.ticks.x = element_blank(),
+        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
+        plot.title = element_text(size = 14),
+        legend.title = element_blank(),
+        plot.tag.position = "bottomleft")
+figure_3.2
 ```
 
-Figure 3 includes countries that experience at least 100 cases of MPX.
+In our dataset, 9 countries were missing population data for age stratums which left us with 20 countries with MPX cases ranging 4 - 2889 and a median of 42.
+Figure 3 includes two plots dividing countries that experienced at least 100 cases of MPX with those that had less than 100 cases.
 The figure parameters include stacked bar graphs that visually represent the age brackets from which a positive case is observed.
-The 30-49 age range is observed to have more positive cases across each country than any of the other age brackets.
-Given the consistency within this observation, the investigators may find that countries with populations with thriving numbers of 30-49-year-old's are affected more by MPX than regions that do not.
-Moreover, the countries listed in this figure are predominantly within the Western Region of Europe; Western Europe includes six countries.
+When comparing the two graphs, the proportion among each age stratum appear similar with the stratum of age < 15 ranging from 9.68%-14.42% and 9.12% - 12.09% in countries with >= 100 cases and countries with < 100 cases, respectively.
+Similar proportions were also seen for age stratum > 64.
+The 30-49 age range makes up the largest proportion of population among the 20 European country in which a MPX case was reported and age stratum information was available.
+In a seperate analyses of countries with confirmed MPX and sex (M/F), no relationship could be seen with country level sex breakdown and number of MPX cases.
 
@@
 
-Through analysis of Figure 3, we are not confident of the initial hypothesis regarding younger (under 30 years old) and older (over 64 years old) populations proportions of country demographics causing a higher risk of infection.
-Review of the demographics associated to Figure 3 provide some incite as to the population age bracket proportions for those countries listed (produced over 100 confirmed cases).
-We observe that the age group for 30-64-year-olds (broken into two categories of 30-49 and 50-64 years) attributes to over 50 % of the population proportion for all six countries.
-When we reviewed the demographic characteristics associated to sex of the populations within each region, we felt that the use of age within our figure provided a better story into potential factors that are causing the occurrence of cases of MPX in Europe.
+Through analysis of Figure 3, we are not confident of the initial hypothesis regarding younger (under 15 years old) and older (over 64 years old) populations proportions of country demographics causing a higher risk of infection.
+Review of the demographics associated to Figure 3 led to an inconclusive relationship between age and MPX.
 
-Given the analysis of this information, health department leadership should begin educational programs and outreach focusing on those 30-64 aged individuals.
+Given the analysis of this information, health department leadership should begin educational programs and outreach focusing on countries in Western and Southern Europe.
